\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-04-25 doc/incertae_sedis.tex
Reformat and tiny edits.

The supertree algorithm of \citet{redelings2017supertree} takes as
input a ranked list $\phyloinputs=\left\{ T_{1},\ldots,T_{n}\right\} $
of leaf-labeled rooted input phylogenies and a single rooted taxonomy
tree $\taxonomy_{p}$ that is ranked below all of the input trees.
As
discussed in \citet{redelings2017supertree}, the preferred tree would
display as many of the highly ranked input splits as possible.
If
computational barriers were not an issue, the method would act by
first producing a ranked list of rooted splits:
$\mathcal{S}=S(T_{1})+S(T_{2})+\ldots+S(T_{n})+S(\mathbb{T})$ where
$S(T_{i})$ denotes a list of non-trivial rooted splits created from a
post-order traversal of tree $i$ and '$+$' denotes concatenation.
Then
the supertree would be produced in a greedy manner by iterating over
the list $S$, and adding each rooted split to a growing rooted
supertree if that split is compatible with the supertree.
In other
words, the set $C$ is initialized to the empty set, and we then
iteratively consider each split in the list $\mathcal{S}$ and insert
it into $C$ if the resulting set remains jointly compatible.
The BUILD
algorithm \citep{AhoSSU1981} is used both to check compatibility.
 and to construct many aspects of a compatible tree from the final set $C$.
Refer to \citet{redelings2017supertree} for details of the procedure, including
a description of an optimization from
  pruning taxa that do not occur in any phylogenetic estimates in pre-processing
  and then reattaching them to a backbone tree.
This optimization is not crucial to the current work, but it does
  cause some groups to attach more tipward than they would if the entire
  tree were constructed using BUILD.

@@

The taxonomy tree, $\taxonomy_{p}$, is derived from OTT.
As mentioned
above, OTT not only contains a taxonomic hierarchy, but also contains
taxonomic labels and annotations (referred to as ``flags'') on the
taxa to carry extra information.
The entire hierarchy encoded by OTT is
a complete taxonomic tree, $\taxonomy^{\ast}.$ Nodes in
$\taxonomy^{\ast}$ have a one-to-one correspondence with a set of
labels $\mathcal{N}$ that are called taxon names.
We will decompose
$\mathcal{N}$ into the taxon names for leaf taxa \emph{$\mathcal{L}$
}and higher taxa $\mathcal{H}$.
We will use $\mathcal{U}$ to refer to
the subset ($\mathcal{U}\subseteq\mathcal{N}$) of taxa that are
annotated with the \emph{incertae sedis }property.
Other annotations
identify dubious taxa and taxonomic names that are unwanted artifacts
of the process of building OTT.
Thus, the taxonomic tree which is used
as an input to the supertree algorithm is pruned version of
$\taxonomy^{\ast}$.
For our purposes, the relevant information from
the taxonomy consists of the taxonomy tree $\taxonomy^{\ast}$, the
labels $\mathcal{N}$, and set $\mathcal{U}$ of \emph{incertae sedis} taxa.
In previous work, taxa in $\mathcal{U}$ and all of their descendants 
were all pruned from
the taxonomic hierarchy when $\taxonomy_{p}$ is extracted from OTT.
In
the present work, we consider supertree approaches that operate on a
taxonomic tree $\taxonomy$, which is produced from
$\taxonomy^{\ast}$but does not prune \emph{incertae sedis} taxa.

>>>>

The supertree algorithm of \citet{redelings2017supertree} takes as input a
ranked list $\phyloinputs=\left\{ T_{1},\ldots,T_{n}\right\} $ of leaf-labeled rooted input phylogenies and a
single rooted taxonomy tree $\taxonomy_{p}$ that is ranked below all of the input trees.
As discussed in \citet{redelings2017supertree}, the preferred tree would
display as many of the highly ranked input splits as possible. If computational
barriers were not an issue, the method would act by first producing a ranked
list of rooted splits: $\mathcal{S}=S(T_{1})+S(T_{2})+\ldots+S(T_{n})+S(\mathbb{T})$ where $S(T_{i})$ denotes a list
of non-trivial rooted splits created from a post-order traversal of tree $i$ and '$+$'
denotes concatenation. Then the supertree would be produced in a greedy manner
by iterating over the list $S$, and adding each rooted split to a growing rooted
supertree if that split is compatible with the supertree. In other words, the set
$C$ is initialized to the empty set, and we then iteratively consider each split in
the list $\mathcal{S}$ and insert it into $C$ if the resulting set remains jointly compatible.
The BUILD algorithm \citep{AhoSSU1981} is used both to check compatibility,  and
to construct many aspects of a compatible tree from the final set $C$.  Refer to
\citet{redelings2017supertree} for details of the procedure, including a
description of an optimization for pruning taxa that do not occur in any
phylogenetic estimates during pre-processing and then reattaching them to a
backbone tree.  This optimization is not crucial to the current work, but it
does cause some groups to attach more tipward than they would if the entire tree
were constructed using BUILD.

@@

The taxonomy tree, $\taxonomy_{p}$, is derived from OTT. As mentioned above, OTT not only
contains a taxonomic hierarchy, but also contains taxonomic labels and
annotations (referred to as ``flags'') on the taxa to carry extra information.
The entire hierarchy encoded by OTT is a complete taxonomic tree, $\taxonomy^{\ast}.$ Nodes in
$\taxonomy^{\ast}$ have a one-to-one correspondence with a set of labels $\mathcal{N}$ that are called
taxon names. We will decompose $\mathcal{N}$ into the taxon names for leaf taxa $\mathcal{L}$ and
higher taxa $\mathcal{H}$. We will use $\mathcal{U}$ to refer to the subset ($\mathcal{U}\subseteq\mathcal{N}$) of taxa that are
annotated with the \emph{incertae sedis }property. Other annotations identify
dubious taxa and taxonomic names that are unwanted artifacts of the process of
building OTT. Thus, the taxonomic tree which is used as an input to the
supertree algorithm is pruned version of $\taxonomy^{\ast}$. For our purposes, the relevant
information from the taxonomy consists of the taxonomy tree $\taxonomy^{\ast}$, the labels $\mathcal{N}$,
and set $\mathcal{U}$ of \emph{incertae sedis} taxa. In previous work, taxa in $\mathcal{U}$ and all
of their descendants were all pruned from the taxonomic hierarchy when $\taxonomy_{p}$ is
extracted from OTT. In the present work, we consider supertree approaches that
operate on a taxonomic tree $\taxonomy$, which is produced from $\taxonomy^{\ast}$but does not prune
\emph{incertae sedis} taxa.
